Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-03-27 ms.tex
added three-panel 100 GHz low-order correction summary figure
outline
added 2nd table
clean-up
started re-write of relative zero level subsection

\subsubsection{Absolute Zero Level}
In \cite{planckdust}, the absolute zero level of infrared thermal dust emission
was set by requiring that \PLANCK~infrared emission tends to zero when 
H\,\textsc{i} is zero, assuming a linear correlation between these two 
measurements at low column density. However, this approach is somewhat
unsatisfying in that there appear to be different slopes of \PLANCK~857 GHz 
versus H\,\textsc{i} for different ranges of H\,\textsc{i} intensity. In 
particular, \PLANCK~857 GHz appears to ``flatten out'' at very low 
H\,\textsc{i}. More quantitavely, we have found using the LAB H\,\textsc{i} 
data \citep{lab} for $-72$$<$$v_{LSR}$$<$$+25$ that the best-fit slope for 
H\,\textsc{i}$<$70 K km/s is a factor of $\sim$1.9 lower than the best fit 
slope for 110 K km/s $<$H\,\textsc{i}$<$200 K km/s, and as a result the implied
zero-point offsets for \PLANCK~857 GHz differ by $\sim$0.37 MJy/sr.

% CITE KALBERLA LAB PAPER !!!!!

@@

Because of these ambiguities in the relationship between 857 GHz and 
H\,\textsc{i} emission, we decided to instead constrain the 857 GHz zero level
by comparison to the FDS99 857 GHz prediction. This renders the \PLANCK~857 GHz
absolute zero level tied largely to the reliable FIRAS zero point, and to some 
lesser extent to H\,\textsc{i} through the \verb|i100| zero point. 
Specifically, we fit a linear model to the FDS99 predicted 857 GHz values as a 
function of \PLANCK~857 GHz. We used a version of the \PLANCK~857 GHz map with 
zodiacal light and point sources removed and smoothed to 1$^{\circ}$ 
resolution, which we will refer to as $\mathcal{I}_{857}$. For this analysis, 
we considered $\mathcal{I}_{857}$ to be the independent variable, as it has 
much higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not a monochromatic 
prediction, as we have incorporated the finite bandpass correction factors 
described in $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine
the dust spectrum shape. We then rebin to $N_{side}$=64 and restrict to pixels 
with $\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to 
degree resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure XX shows a scatterplot of 
$\mathcal{I}_{857}$ versus $\mathcal{F}_{857}$, with a moving median and linear
fit overplotted. The linear fit was performed with uniform weights and 
iterative outlier rejection. The best-fit linear model is given by 
$\mathcal{F}_{857}$=0.994$\mathcal{I}_{857}$$-$0.018. It is encouraging that 
the slope is quite close to unity, and the offset near zero.

The formal statistical errors on these parameters are small. Just as in the 
case of performing the regression versus  H\,\textsc{i}, 
systematic issues will dominate the uncertainty in the derived zero point
offset. The likely culprits in this case are imperfect zodiacal light 
corrections and the FDS99 temperature map. To quantify the level of systematic
error, we split the sky into four quadrants, with the dividing boundaries
being $b$=0$^{\circ}$ and $l$=0$^{\circ}$, $l$=180$^{\circ}$. We again
restricted to $\mathcal{I}_{857}$$<$2.15 MJy/sr, and repeated the 
regression in each quadrant. The slopes varied from 0.948 to 0.993, with an
rms of 0.0188, while the offsets varied from $-$0.056 MJy/sr to $+$0.089 
MJy/sr, with an rms of 0.0586 MJy/sr.

@@

% in previous sentence fill in specifics of slope and offset

% sentence saying slope is close to unity, and relationship indeed looks linear

% sentence or two explaining split into quadrants, differences in slope
% and offset probably due to zodiacal light and/or issues with FDS T2 map
% say how this quadrants exercise (via rms) allows us to define 857 GHz
% uncertainty

@@

\begin{deluxetable*}{rrrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{6} 
\tablewidth{0pc} 
\tablecaption{\label{obs} Planck/IRAS Map Properties \& Pre-processing} 
\tablehead{
\colhead{$\nu$ (GHz)} &
\colhead{Offset ($K_{CMB}$)} & 
\colhead{Dipole ($K_{CMB}$)} & 
\colhead{Quadrupole ($K_{CMB}$)} & 
\colhead{$c_{\nu}$} &
\colhead{FWHM ($'$)}
}
\startdata
100 & 1.69$\times$10$^{-5}$$\pm$3.61$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & $-$7.17$\times$10$^{-6}$ &  0.0054 & 9.66 \\
143 & 3.58$\times$10$^{-5}$$\pm$7.58$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & - & 0.0054 &  7.27 \\
217 & 7.79$\times$10$^{-5}$$\pm$2.60$\times$10$^{-6}$ & $-$1.40$\times$10$^{-5}$ & - & 0.0054 &  5.01 \\
353 & 2.76$\times$10$^{-4}$$\pm$1.95$\times$10$^{-5}$ & $-$3.08$\times$10$^{-5}$ & - & 0.012 & 4.86 \\
    & Offset (MJy/sr) & Dipole (MJy/sr) & Quadrupole (MJy/sr) & & \\ \cline{2-4} \\ [-2ex]
545 & 7.28$\times$10$^{-2}$$\pm$1.99$\times$10$^{-2}$ & 1.63$\times$10$^{-2}$  & - & 0.10 & 4.84 \\
857 & 1.82$\times$10$^{-2}$$\pm$6.02$\times$10$^{-2}$ &  - & - & 0.10 & 4.63 \\
3000 & 0.0$\pm$4.3$\times$10$^{-2}$ & - & - & 0.10 & 6.1
\enddata
\end{deluxetable*}

@@

At 353 GHz, 545 GHz, 857 GHz, we have 
adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
 by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
then by correlating the other \PLANCK~channels against the corrected 857 GHz 
map. We have not altered the \verb|i100| zero-point in any way. We have also 
removed the Solar dipole residual according to \cite{planckdust}, by 
subtracting a dipole pattern oriented toward 
$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
Table 1.

\begin{figure}
\begin{center}
\epsfig{file=scatter_857_217.eps, width=3.4in}
\caption{\label{fig:scatter} Scatter plot of \PLANCK~217 GHz versus \PLANCK~857
GHz. Left: before correcting for offset and residual Solar dipole. Right: after
correcting or the best-fit offset and dipole amplitude, as described in 
$\S$\ref{sec:zp}. The dashed red line shows the best-fit linear relationship
between 857 GHz and 217 GHz.}
\end{center}
\end{figure}

The offset and residual dipole parameters for \PLANCK~217 GHz were not 
available from \cite{planckdust} Table 1. In order to set the zero-point of 
Galactic emission in the 217 GHz map, we simultaneously fit the constant 
offset and residual dipole amplitude. Restricting to regions of low column 
density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ 
km\,s$^{-1}$), we modeled the 217 GHz map as:

\begin{equation}
\mathcal{I}_{217, i} = a\mathcal{I}_{857, i} + b\mathcal{D}_{i} + c
\end{equation}

With $i$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low H\,\textsc{i} cut, we have an overconstrained linear model with
three parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improved correlation striking in this scatterplot, 
but the residual dipole also appears visually well-corrected in the 217 GHz 
map itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

@@

@@

\subsection{Validation with FIRAS}

>>>>

\subsubsection{Absolute Zero Level}
\label{sec:zp_abs}
In \cite{planckdust}, the absolute zero level of infrared thermal dust emission
was set by requiring that \PLANCK~infrared emission tends to zero when 
H\,\textsc{i} is zero, assuming a linear correlation between these two 
measurements at low column density. However, this approach is somewhat
unsatisfying in that there appear to be different slopes of \PLANCK~857 GHz 
versus H\,\textsc{i} for different ranges of H\,\textsc{i} intensity. In 
particular, \PLANCK~857 GHz appears to ``flatten out'' at very low 
H\,\textsc{i}. More quantitavely, we have found using the LAB H\,\textsc{i} 
data \citep{lab} for $-72$$<$$v_{LSR}$$<$$+25$ that the best-fit slope for 
H\,\textsc{i}$<$70 K km/s is a factor of $\sim$1.9 lower than the best fit 
slope for 110 K km/s $<$H\,\textsc{i}$<$200 K km/s, and as a result the implied
zero-point offsets for \PLANCK~857 GHz differ by $\sim$0.37 MJy/sr.

@@

\begin{figure}
\begin{center}
\epsfig{file=dipole_all.eps, width=3.4in}
\caption{\label{fig:dip}  Scatter plots of \PLANCK~143, 217, 353, and 545 GHz 
versus \PLANCK~857 GHz. Left: before correcting for offset and residual Solar 
dipole. Right: after correcting or the best-fit offset and dipole amplitude, as
described in $\S$\ref{sec:zp}. The dashed red line shows the best-fit linear 
relationship in each case.}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
\epsfig{file=quad_100.eps, width=6.5in}
\caption{\label{fig:quad} Summary of low-order corrections at 100 GHz.}
\end{center}
\end{figure*}

Because of these ambiguities in the relationship between 857 GHz and 
H\,\textsc{i} emission, we decided to instead constrain the 857 GHz zero level
by comparison to the FDS99 857 GHz prediction. This renders the \PLANCK~857 GHz
absolute zero level tied largely to the reliable FIRAS zero point, and to some 
lesser extent to H\,\textsc{i} through the \verb|i100| zero point. 
Specifically, we fit a linear model to the FDS99 predicted 857 GHz values as a 
function of \PLANCK~857 GHz. We used a version of the \PLANCK~857 GHz map with 
zodiacal light and point sources removed and smoothed to 1$^{\circ}$ 
resolution, which we will refer to as $\mathcal{I}_{857}$. For this analysis, 
we considered $\mathcal{I}_{857}$ to be the independent variable, as it has 
much higher S/N than the FDS99 prediction, henceforward referred to as 
$\mathcal{F}_{857}$. Note that $\mathcal{F}_{857}$ is not a monochromatic 
prediction, as we have incorporated the finite bandpass correction factors 
described in $\S$\ref{sec:bpcorr}, using the FDS99 temperature map to determine
the dust spectrum shape. We then rebin to $N_{side}$=64 and restrict to pixels 
with $\mathcal{I}_{857}$$<$2.15 MJy/sr. Since \PLANCK~857 GHz smoothed to 
degree resolution has very high S/N, we can safely perform such a cut on 
$\mathcal{I}_{857}$. Figure \ref{fig:fdsref} shows a scatterplot of 
$\mathcal{I}_{857}$ versus $\mathcal{F}_{857}$, with a moving median and linear
fit overplotted. The linear fit was performed with uniform weights and 
iterative outlier rejection. The best-fit linear model is given by 
$\mathcal{F}_{857}$=0.994$\mathcal{I}_{857}$$-$0.018. It is encouraging that 
the slope is quite close to unity, and the offset near zero.

The formal statistical errors on these parameters are small. Just as in the 
case of performing the regression versus  H\,\textsc{i}, systematic problems 
will dominate the uncertainty in the derived zero point offset. The systematics
likely to dominate our FDS-based zero level are imperfect zodiacal light 
corrections and the FDS99 temperature map. To quantify the level of systematic 
error, we split the sky into four quadrants, with the dividing boundaries being
 $b$=0$^{\circ}$ and $l$=0$^{\circ}$, $l$=180$^{\circ}$. We again restricted to
 $\mathcal{I}_{857}$$<$2.15 MJy/sr, and repeated the regression in each 
quadrant. The slopes varied from 0.948 to 0.993, with an rms of 0.0188, while 
the offsets varied from $-$0.056 MJy/sr to $+$0.089 MJy/sr, with an rms of 
0.0586 MJy/sr.

@@

@@

\begin{deluxetable*}{rrrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{6} 
\tablewidth{0pc} 
\tablecaption{\label{table:offs} Planck/IRAS Map Properties \& Pre-processing} 
\tablehead{
\colhead{$\nu$ (GHz)} &
\colhead{Offset ($K_{CMB}$)} & 
\colhead{Dipole ($K_{CMB}$)} & 
\colhead{Quadrupole ($K_{CMB}$)} & 
\colhead{$c_{\nu}$} &
\colhead{FWHM ($'$)}
}
\startdata
100 & 1.69$\times$10$^{-5}$$\pm$3.61$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & $-$7.17$\times$10$^{-6}$ &  0.0054 & 9.66 \\
143 & 3.58$\times$10$^{-5}$$\pm$7.58$\times$10$^{-7}$ & $-$1.08$\times$10$^{-5}$ & - & 0.0054 &  7.27 \\
217 & 7.79$\times$10$^{-5}$$\pm$2.60$\times$10$^{-6}$ & $-$1.40$\times$10$^{-5}$ & - & 0.0054 &  5.01 \\
353 & 2.76$\times$10$^{-4}$$\pm$1.95$\times$10$^{-5}$ & $-$3.08$\times$10$^{-5}$ & - & 0.012 & 4.86 \\
    & Offset (MJy/sr) & Dipole (MJy/sr) & Quadrupole (MJy/sr) & & \\ \cline{2-4} \\ [-2ex]
545 & 7.28$\times$10$^{-2}$$\pm$1.99$\times$10$^{-2}$ & 1.63$\times$10$^{-2}$  & - & 0.10 & 4.84 \\
857 & 1.82$\times$10$^{-2}$$\pm$6.02$\times$10$^{-2}$ &  - & - & 0.10 & 4.63 \\
3000 & 0.0$\pm$4.3$\times$10$^{-2}$ & - & - & 0.10 & 6.1
\enddata
\end{deluxetable*}

@@

In the course of this study we use not only \PLANCK~857 GHz, but also all of
the remaining \PLANCK~HFI bands as well as \verb|i100|. To derive the 
absolute zero level offsets that must be applied to each of the five
lowest-frequency \PLANCK~bands, we perform a regression versus the 
\PLANCK~857 GHz map corrected for the absolute zero level offset derived in
 $\S$\ref{sec:zp_abs}. We assume no offset need be applied
to \verb|i100|, which has a zero level tied to H\,\textsc{i} by SFD.
% maybe need to restate the idea that we're deriving zero points by assuming
% that the band of interest must approach zero linearly as the reference
% (in this case 857 ghz) approaches zero.

Inspecting the HFI maps at 100-545 GHz, the need for low-order corrections beyond
a scalar offset were evident. In particular, we noticed the presence of a residual 
Solar dipole. To best recover the amplitudes of the low-order corrections, it is
necessary to have the highest-possible signal in the independent variable of our
regression. For this reason we have used 857 GHz as a reference for the other HFI 
bands, as opposed to the FDS99 predictions or H\,\textsc{i}.

Importantly, we have performed one regression per HFI band to simultaneously fit for the 
zero level offset, the slope relative to 857 GHz, and the residual dipole amplitude.
For each 100-545 GHz HFI band, we restrict tor regions of low column density
(H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ km\,s$^{-1}$)
and fit the following model:

%At 353 GHz, 545 GHz, 857 GHz, we have 
%adopted the zero-point offsets of \cite{planckdust} Table 1, which were derived
% by correlating \PLANCK~857 GHz versus H\,\textsc{i} at low column density, and
%then by correlating the other \PLANCK~channels against the corrected 857 GHz 
%map. We have not altered the \verb|i100| zero-point in any way. We have also 
%removed the Solar dipole residual according to \cite{planckdust}, by 
%subtracting a dipole pattern oriented toward 
%$(l, b) = (263.99^{\circ}, 48.26^{\circ})$ and scaled according to their 
%Table \ref{table:offs}.

%The offset and residual dipole parameters for \PLANCK~217 GHz were not 
%available from \cite{planckdust} Table 1. In order to set the zero-point of 
%Galactic emission in the 217 GHz map, we simultaneously fit the constant 
%offset and residual dipole amplitude. Restricting to regions of low column 
%density (H\,\textsc{i} $<$ 200 K\,km\,s$^{-1}$ for $-72$$<$$v_{LSR}$$<$$+25$ 
%km\,s$^{-1}$), we modeled the 217 GHz map as:

\begin{equation}
\mathcal{I}_{\nu_i, p} = a\mathcal{I}_{857, p} + b\mathcal{D}_{p} + c
\end{equation}

With $p$ denoting a single nside=64 HEALPix pixel in the maps 
$\mathcal{I}_{857}$, $\mathcal{I}_{217}$, and $\mathcal{D}$. Here 
$\mathcal{I}_{857}$ is the \PLANCK~857 GHz map with zodiacal emission, compact 
sources, and constant offset removed, smoothed to $1^{\circ}$ resolution. 
$\mathcal{I}_{217}$ is the corresponding $1^{\circ}$ resolution \PLANCK~217 
GHz map with zodiacal emission and compact sources removed. $\mathcal{D}$ is a
scaling of the Solar dipole with unit amplitude. Because $\sim$18,000 pixels
satisfy the low H\,\textsc{i} cut, we have an overconstrained linear model with
three parameters: a, b, and c. $a$ represents the slope of \PLANCK~217 GHz 
versus \PLANCK~857 GHz assuming they are linearly correlated, and is not
of any importance to our downstream analysis. $b$ represents the scaling of 
the Solar dipole that makes the 857 GHz vs. 217 GHz correlation tightest, and 
$c$ represents the constant offset necessary to make the zero level of the 217 
GHz map consistent with that of the 857 GHz map. Performing a linear least
squares fit with uniform weights and iterative outlier rejection, we find 
$a=4.33\times10^{-5}\ K/(MJy/sr)$, $b=-1.40\times10^{-5}\ K$, and 
$c=8.12\times10^{-5}\ K$. Note that $b$ represents 0.42\% of the Solar dipole 
amplitude, which is within the calibration uncertainty of the \PLANCK~217 GHz 
band. Figure \ref{fig:scatter} shows scatterplots of the 857 GHz versus 217 
GHz correlation before and after correcting for our best-fit offset and dipole 
residual. Not only is the improved correlation striking in this scatterplot, 
but the residual dipole also appears visually well-corrected in the 217 GHz 
map itself. Before performing thermal dust fits, we therefore subtract $c$ and 
$b\mathcal{D}$ from the 217 GHz map.

%in above paragraph, clarify what dipole means i.e. in what direction
% on sky it's oriented

%mention the special case of 100 GHz w/ quadrupole term ... do i need
% another equation ?

% mention how offset/dipole values compare to those derived in the Planck
% team paper ?

@@

\section{Global Model Parameters}

%%%%

\begin{deluxetable}{rrrrr} 
\tabletypesize{\scriptsize}
\tablecolumns{5} 
\tablewidth{0pc} 
\tablecaption{\label{obs} Global Model Parameters} 
\tablehead{
\colhead{Data} &
\colhead{$f_1$} &
\colhead{$q_1/q_2$} & 
\colhead{$\beta_1$} & 
\colhead{$\beta_2$}
}
\startdata
FIRAS+\IRAS & 0.0363 & 13.0 & 1.67 & 2.70 \\
\PLANCK+\IRAS & 0.0267 & 3.68 & 1.19 & 2.61
\enddata
\end{deluxetable}

%%%%

\subsection{Correlation Slopes}

% multi-panel figure showing more scatterplots of Planck vs. FDS predictions

\subsection{Hierarchical Approach?}

@@

%\subsection{Validation with FIRAS}
